---
title: "Lab 5"
author: "Sean Kim"
format:
  html:
    embed-resources: true
---

```{r}
library(data.table)
library(dtplyr)
library(dplyr)
```

```{r}
met <- fread("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz")
met[, elev := fifelse(elev ==9999.0, NA_integer_, elev)]
met[, temp := fifelse(temp < -15, NA_integer_, temp)]

met <- met[!is.na(elev) & !is.na(temp)]
```

```{r}
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

stations[, USAF := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE := fifelse(STATE == "", NA_character_, STATE)]

stations <- unique(stations[, list(USAF, CTRY, STATE)])

stations <- stations[!is.na(USAF)]

stations[, n:= 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

```{r}
alt_merge <- met %>%
  inner_join(stations, by = c("USAFID" = "USAF"))
alt_merge <- alt_merge[!is.na(lat)]
```

Question 1: Representative station for the US

```{r}
quantiles <- alt_merge %>%
  summarize(
    MedianTemp = quantile(temp, probs = 0.5, na.rm = TRUE),
    MedianWindSpeed = quantile(wind.sp, probs = 0.5, na.rm = TRUE),
    MedianPressure = quantile(atm.press, probs = 0.5, na.rm = TRUE)
  )
```

Median Stations representative of the US:

Temp: 23.5C

Wind speed: 2.1

Pressure: 1014

```{r}
closest_stations <- alt_merge[, ClosePress := abs(atm.press - quantiles$MedianPressure)]
closest_stations <- alt_merge[, CloseTemp := abs(temp - quantiles$MedianTemp)]
closest_stations <- alt_merge[, CloseWind := abs(wind.sp - quantiles$MedianWindSpeed)]
                            

representative_stations_US <- closest_stations[, .SD[which.min(CloseTemp + ClosePress + CloseWind),
                                                .(CloseTemp, ClosePress, CloseWind)],
                                               by = USAFID]
representative_stations_US
```

The three stations that best represent the US in terms of median temp, wind speed, and atmospheric pressure are USAFID 690150, 720175, and 720198.

Question 2: Representative Station per State

```{r}
MedianState <- alt_merge[, .(
    MedianTemp      = median(temp, na.rm=TRUE),
    MedianWindSpeed   = median(wind.sp, na.rm=TRUE),
    MedianPressure = median(atm.press, na.rm = TRUE), 
    Latitude = median(lat, na.rm = TRUE), 
    ID = unique(USAFID)
    ),
    by = STATE
    ][order(STATE)]
MedianState[, EuclideanDist := sqrt((MedianTemp - median(MedianTemp))^2 +
                                    (MedianWindSpeed - median(MedianWindSpeed))^2 +
                                    (MedianPressure - median(MedianPressure))^2)]
representative_stations <- MedianState[, .SD[which.min(abs(MedianTemp) + 
                                                  abs(MedianWindSpeed) + 
                                                  abs(MedianPressure)), 
                                                .(ID, Latitude, MedianTemp, MedianWindSpeed, MedianPressure)],
                                         by = STATE]
representative_stations
```

Displayed table of the sites (ID = USAFID) that best represent their respective states.

Question 3 - In the middle?

```{r}
library(leaflet)
```

```{r}
state_midpoints <- alt_merge[, .(MidLat = mean(lat, na.rm = TRUE),
                                   MidLon = mean(lon, na.rm = TRUE)),
                               by = STATE]
alt_merge[, MidPointLat := mean(lat, na.rm = TRUE), by = STATE]
alt_merge[, MidPointLon := mean(lon, na.rm = TRUE), by = STATE]
alt_merge[, DistanceToMidpoint := sqrt((lat - MidPointLat)^2 + (lon - MidPointLon)^2)]

closest_stations_midpoint <- alt_merge[, .SD[which.min(DistanceToMidpoint)],
                                         by = STATE]


map <- leaflet() %>%
  addTiles()

map <- map %>%
  addCircleMarkers(
    data = state_midpoints,
    lat = ~MidLat,
    lng = ~MidLon,
    radius = 3,
    color = "blue",
    popup = ~STATE
  )

# Plot stations identified in this question with a different color
map <- map %>%
  addCircleMarkers(
    data = closest_stations_midpoint,
    lat = ~lat,
    lng = ~lon,
    radius = 3,
    color = "red",
    popup = ~USAFID
  )

# Display the map
map
```

Question 4 - Means of means

```{r}
met_means <- alt_merge [, .(avgTemp = mean(temp, na.rm = TRUE)), by = STATE]

avgTempCat <- alt_merge [, avgTemp := mean(temp, na.rm = TRUE), by = STATE]
avgTempCat[, TempCat := cut(avgTemp, breaks = c(-Inf, 20, 25, Inf), labels = c("Low", "Mid", "High"))]

summary_table <- alt_merge[, .(NumEntries = .N,
                                 NumNAEntries = sum(is.na(temp)),
                                 NumStations = length(unique(USAFID)),
                                 NumStates = length(unique(STATE)),
                                 MeanTemp = mean(temp, na.rm = TRUE),
                                 MeanWindSpeed = mean(wind.sp, na.rm = TRUE),
                                 MeanPressure = mean(atm.press, na.rm = TRUE)),
                             by = TempCat]
summary_table
```
